### 2020-04-13 dataclean.tex
miscellaneous ...
mostly km model revamped

A typical GPU data record produced by the methods in the last section
is of the following form
{\small
\begin{verbatim}
0323812007945 | c17-4c1s3n1 | 01/10/2013 02:54:45 | 02/02/2013 11:32:29
 | c13-1c1s3n3 | 01/21/2014 21:10:42 | 08/01/2017 02:15:02
 | c0-1c1s3n3 | 10/11/2013 15:57:33 | 10/12/2013 22:09:31
 | c21-1c2s5n0 | 03/19/2013 15:48:11 | 05/29/2013 11:54:11
0325216047736 | c18-4c1s5n1 | 03/07/2017 02:15:01 | 03/12/2019 03:19:11
0323812008856 | c5-4c0s7n0 | 01/10/2013 02:54:45 | 01/25/2013 15:29:58
 | c0-6c1s7n2 | 10/21/2013 14:28:19 | 10/28/2013 17:52:44
 | c3-3c1s5n0 | 05/29/2013 11:54:11 | 05/29/2013 11:54:11
 | c23-6c1s7n2 | 01/21/2014 21:10:42 | 11/02/2018 14:42:34
 | | DBE | 11/02/2018 14:42:34
\end{verbatim}
}
\noindent where records for three GPUs are shown. Each record starts
with a serial number and locations are coded with {\tt c{\it
    col-row}c{\it cage}s{\it slot}n{\it node}}. The first GPU record
shows installation in locations {\tt c17-4c1s3n1}, {\tt c21-1c2s5n0},
{\tt c0-1c1s3n3}, with periods off the system, and finally in {\tt
  c13-1c1s3n3}, where it stays until August 1, 2017, after which it is
not seen again. The second is intalled in location {\tt c18-4c1s5n1},
where it stays until the last data collection date on March 3, 2019,
at 3:19:11 AM. The third GPU is first installed in locations {\tt
  c5-4c0s7n0}, {\tt c3-3c1s5n0},{\tt c0-6c1s7n2}, {\tt c23-6c1s7n2},
where a ``DBE'' is observed on November 2, 2018, and it is not seen
again.

@@

A plot of the life of 100 random GPUs on Titan after cleaning the data:
\begin{description}
\item[starts:] black dots (gray for zero lifetime)
\item[lifetimes:] black lines
\item[DBEs:] red triangles
\item[Off The BUS:] blue triangles
\item[Last seen:] black ]
\item[overlaps:] red square on left when they exist
\end{description}

\begin{figure}[ht]
  \includegraphics[width=6in]{sn.pdf}
  \caption{GPU serial number view of life and failures.}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=6in]{location.pdf}
  \caption{GPU location view of life and failures.}
\end{figure}

>>>>

@@

To get some intuition for the GPU lifetimes on Titan, we give two views of
a hundred GPUs (Figure~\ref{fig:gpuview}) and a hundred locations
(Figure~\ref{fig:locview}). The GPU view is ``by serial number'' that
shows when each GPU was installed and removed at various locations as
well as its DBE and OTB events. The second view by location, shows
when different GPUs were installed and removed at a location, and
their OTB and DBE events.
\begin{figure}[ht]
  \includegraphics[width=6in]{figs/sample_sn.pdf}
  \caption{GPU serial number view of life and failures. Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:gpuview}
\end{figure}
\begin{figure}[ht]
  \includegraphics[width=6in]{figs/sample_loc.pdf}
  \caption{GPU location view of life and failures.  Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:locview}
\end{figure}
The serial number view clearly shows the two batches and the more
frequent OTB and DBE events in the old batch. It is also clear from
this view that most new GPUs stay at the initial install location
whereas the old GPUs were occasionally reinstalled at new locations.
The location view shows that each location was operational almost all
the time with small gaps when GPUs were changed out.

\fix{Were the GPUs swapped out as a blade or were individual GPUs
  swapped on a blade?}

### 2020-04-13 dataprep.tex
mostly km model revamped

\section{Data Collection and Preprocessing}
\label{sec:data}
\fix{Don's section}

>>>>

\section{Data Collection and Preprocessing}
\label{sec:data}
\fix{Don's section}
Data on GPU life times is constructed from two sources: inventory runs
and failure event records. Two types of failure events were collected:
Double Bit Error (DBE) and Off the Bus (OTB). DBE was found to be the
``signature'' error of eventual GPU failures.

An inventory of GPU serial numbers and their locations is recorded
each time the system boots. \fix{Don, you mentioned that this was more
regular later - daily?} This typically occurs once every few days but
can be as frequent as a few times per day. An inventory at boot time
typically occurs over a span of \fix{??} minutes. A separate file is
produced at each boot time.

The DBE and OTB events are also recorded separately from boot
time inventories. This is recorded as . . . files.

Inventories are taken at boot time, which can occur anywhere from
several times per day to more than a month
apart. Figure~\ref{fig:inventory} shows inter-inventory times that
appear in the data.
\begin{figure}[bth]
  \begin{center}
    \includegraphics[width=\textwidth]{figs/attention_intervals001.pdf}
  \end{center}
  \caption{Inter-inventory times, obrained by taking intervals between
    unique dates in the data. Note square root scale on $y$-axis.}
  \label{fig:inventory}
\end{figure}
\fix{George: Add color for another variable? (new-old, date)??}
Notably, there was a period of over a month that Titan operated
without a single reboot.

Initial processing updates an inventory by checking the Serial Number
(SN) and Location of each GPU and creating or updating a separate
record for each contiguously observed SN-Location combination.

A typical GPU data record is of the form shown in Fig.~\ref{fig:dataraw},
\begin{figure}
{\small
\begin{verbatim}
0323812007945 | c17-4c1s3n1 | 01/10/2013 02:54:45 | 02/02/2013 11:32:29
 | c13-1c1s3n3 | 01/21/2014 21:10:42 | 08/01/2017 02:15:02
 | c0-1c1s3n3 | 10/11/2013 15:57:33 | 10/12/2013 22:09:31
 | c21-1c2s5n0 | 03/19/2013 15:48:11 | 05/29/2013 11:54:11
0325216047736 | c18-4c1s5n1 | 03/07/2017 02:15:01 | 03/12/2019 03:19:11
0323812008856 | c5-4c0s7n0 | 01/10/2013 02:54:45 | 01/25/2013 15:29:58
 | c0-6c1s7n2 | 10/21/2013 14:28:19 | 10/28/2013 17:52:44
 | c3-3c1s5n0 | 05/29/2013 11:54:11 | 05/29/2013 11:54:11
 | c23-6c1s7n2 | 01/21/2014 21:10:42 | 11/02/2018 14:42:34
 | | DBE | 11/02/2018 14:42:34
\end{verbatim}
}
\caption{A few records of raw data produced from inventories that is
  processed further in our analysis.}
\label{fig:dataraw}
\end{figure}
\noindent where records for three GPUs are shown. Each record starts
with a serial number and locations are coded with {\tt c{\it
    col-row}c{\it cage}s{\it slot}n{\it node}}. The first GPU record
shows installation in locations {\tt c17-4c1s3n1}, {\tt c21-1c2s5n0},
{\tt c0-1c1s3n3}, with periods off the system, and finally in {\tt
  c13-1c1s3n3}, where it stays until August 1, 2017, after which it is
not seen again. The second is intalled in location {\tt c18-4c1s5n1},
where it stays until the last data collection date on March 3, 2019,
at 3:19:11 AM. The third GPU is first installed in locations {\tt
  c5-4c0s7n0}, {\tt c3-3c1s5n0},{\tt c0-6c1s7n2}, {\tt c23-6c1s7n2},
where a ``DBE'' is observed on November 2, 2018, and it is not seen
again.

### 2020-04-13 intro.tex
revised assignments
mostly km model revamped

\section{Introduction}
\label{sec:intro}
\fix{Dropped Jim's initial email here. Mostly Jim and Christian}
The Cray XK7 Titan is rapidly nearing its decommissioning date for
DOE/ASCR. The last production runs will be in July of this year, and
then we will stop services, clean up a few things and turn it off.

This system was \#1 in the world for a very long time, and has remained
a critically important computer system. It defied scale with 18,688
individual compute nodes, and delivered tens of billions of computing
hours to DOE mission-critical programs for nearly 7 years.

This was quite an interesting machine from a reliability perspective.
We were forced to execute three very significant rework cycles, two on
the mechanical assembly affecting the PCIe connector from the GPU
daughtercard to the motherboard, and more recently, a replacement of
about 11,000 GPU assemblies because of a failing resistor on a printed
circuit board.

@@

I am looking for help, very specifically with the statistics portion
of this, i.e. what do the failure rates for specific components look
like? What distributions do those failure rates follow? How could
appropriate models extrapolate failure rates beyond the end of service
date (the dotted lines…) ?

@@

>>>>

\section{Introduction}
\label{sec:intro}
The Cray XK7 Titan supercomputer was the \#1 system in the world for a
very long time, and has remained a critically important computer
system through the end of its life in the Summer of 2019. It defied
scale with 18,688 individual compute nodes and delivered tens of
billions of computing hours to the U.S. Oepartment of Energy
mission-critical programs for nearly 7 years.

Titan was quite an interesting machine from a reliability perspective.
Its operation was forced to execute three very significant rework cycles, two on
the mechanical assembly affecting the PCIe connector from the GPU
daughtercard to the motherboard, and a replacement of about 11,000 GPU
assemblies because of a failing resistor on a printed circuit board.

Figure~\ref{fig:chronology} illustrates the chronology of the rework
cycles. It shows the number of GPU changes at periodic inventories.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=\textwidth]{figs/chronology.pdf}
  \caption{Daily volume of GPU swaps with labels on most frequent dates.}
  \label{fig:chronology}
\end{figure}

@@

I am looking for help, very specifically with the statistics portion
of this, i.e. what do the failure rates for specific components look
like? What distributions do those failure rates follow? How could
appropriate models extrapolate failure rates beyond the end of service
date (the dotted lines) ?

@@

\fix{Need a description of Titan GPU events timeline. Jim or Don?}

### 2020-04-13 survival.tex
mostly km model revamped

\section{Survival Analysis}
\label{sec:survival}
\fix{George's survival analysis, check \protect\pkg{finalfit}}

Kaplan-Meier survival analysis gives the probability of survival
beyond a given time. It is a nonparametric technique that makes no
specific failure model assumptions, such as Weibull, Exponential,
etc. It can also split the data into subpopulations and compute
separate survival curves.

If $T$ is taken to be the random variable of a GPU fail time, then its
probability density function is $f(t)$ and its cumulative distribution
function $F(t) = Pr\{T < t\}$ gives the probability that a GPU fails
by duration $t$. The survival function is then 
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t) = \int_t^\infty f(x)dx,
\end{displaymath}
which is the probability of being operational just before duration $t$.

@@

\begin{figure}
  \includegraphics[width=5in]{cox.pdf}
  \caption{GPU hazard ratios from Cox regression model.}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=5in]{glmnet.pdf}
  \caption{GPU variable importance from glmnet penalized regression analysis.}
\end{figure}

>>>>

\section{Survival Analysis}
\label{sec:survival}
 Kaplan-Meier survival analysis (KM) gives the probability of survival
beyond a given time. It is a nonparametric technique that makes no
specific failure model assumptions, such as Weibull, Exponential,
etc. It can also split the data into subpopulations and compute
separate survival curves.

If $T$ is taken to be the random variable of a GPU fail time, then its
probability density function is $f(t)$ and its cumulative distribution
function $F(t) = Pr\{T < t\}$ gives the probability that a GPU fails
by duration $t$. The survival function is then 
\begin{displaymath}
  S(t) = Pr\{T \geq t\} = 1 - F(t) = \int_t^\infty f(x)dx,
\end{displaymath}
which is the probability of being operational at duration
$t$.

We use reliability censoring concepts to use information from units
that were proactively removed before failing as well as those still in
operation till the end. We use the R package {\tt survival} for the KM
analysis.
\begin{figure}
  \centering
  \includegraphics[width=6in]{figs/km_cage-node_o001.pdf}
  \caption{Survival probability from Kaplan-Meier analysis for {\tt
      old} batch GPUs, stratified by {\tt cage} and {\tt node}.}
  \label{fig:km-old-cage-node}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=6in]{figs/km_cage-node_c001.pdf}
  \caption{Here we consider the effect of cage on survival among the
    old batch.}
  \label{fig:km-all-cage-node}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=5in]{figs/cox_o001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt old}
    batch.}
\end{figure}
\begin{figure}
  \includegraphics[width=5in]{figs/cox_n001.pdf}
  \caption{GPU hazard ratios from Cox regression model on {\tt new}
    batch.}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=5in]{figs/glmnet001.pdf}
  \caption{GPU variable importance from glmnet penalized regression analysis.}
\end{figure}
